---
title: "ESS"
author: "Rebecca & Simon "
output: html_notebook
---

## ToDo

* Build github Repo (Name?) -> social data science journey.
* Derive research question
* data wrangling
* fitting models
* write semester paper:
    + Die Studierenden führen für 12 LP abschließend eine eigene Untersuchung durch zu **zwei oder mehr selbst gewählten Beteiligungs- bzw. Entscheidungsprozessen**, um eine konkrete Forschungsfrage zu beantworten (ca. 30 Seiten). Wieviele Seiten zu Zweit?


## Research Question(s)

* internal and external political efficacy  -> likelihood to vote right/populist or green party? Cultural clash. 
* Are citizens who live in systems that provide extensive direct democratic decision-making processes characterized by higher levels of trust in political institutions and higher external political efficacy?
* ESS in combination with direct democracy dataset?
* Macro predictors: polticial system or number of referenda
* Multilevel-Baysian Regression. 



## Variable Selection  



## Packages

```{r}
pacman::p_load(dplyr, ggplot2, readr, haven, broom, purrr, tidyr, magrittr, labelled, sjPlot)

#devtools::install_github("larmarange/labelled")
```

## Data

```{r}
ess <- read_sav("data/ESS8e01.sav")
```

First, inspect the data.

```{r}
dim(ess)
dplyr::glimpse(ess)
```

## binoculaR

`binoculaR` is an own product developed to mimic SPSS in order to quickly inspect variables and their corresponding levels/ attributes. 

```{r}
#devtools::install_github("systats/binoculaR", force = T)
selected <- binoculaR::binoculaR(ess)
save(selected, file = "selected.Rdata")
```

```{r}
load("selected.Rdata")
ess_sub <- ess %>%
  select(selected$index, hinctnta, trstprt, trstplt, edulvlb, -trstun) %>%
  rename(
    id = idno,
    news_pol = nwspol,
    inet = netusoft,
    poli_intr = polintr,
    have_say = psppsgva,
    pol_grp = actrolga,
    have_infl = psppipla, # double?
    in_eff = cptppola,
    trust_parl = trstprl, 
    trust_pol = trstplt,
    trust_party = trstprt,
    trust_leg = trstlgl,
    trust_police = trstplc,
    trust_eu = trstep,
    vote_at = prtvtbat,
    vote_be = prtvtcbe,
    vote_ch = prtvtfch, 
    vote_cz = prtvtdcz, 
    vote_e1 = prtvede1, 
    vote_fi = prtvtdfi, 
    vote_fr = prtvtcfr,
    vote_gb = prtvtbgb,
    vote_ie = prtvtbie,
    vote_il = prtvtcil,
    vote_is = prtvtbis,
    vote_nl = prtvtfnl,
    vote_no = prtvtbno,
    vote_se = prtvtbse,
    vote_si = prtvtesi,
    party_act = wrkprty,
    sign_pet = sgnptit,
    demo = pbldmn,
    lrscale = lrscale,
    s_life = stflife,
    s_econ = stfeco,
    s_gov = stfgov,
    s_dem = stfdem,
    state_edu = stfedu,
    sate_health = stfhlth,
    eu_state = euftf,
    imm_econ = imbgeco,
    imm_same = imsmetn,
    imm_diff = imdfetn,
    imm_poor = impcntr,
    rel = rlgatnd,
    citz = ctzcntr,
    social_lazy = sblazy,
    gndr = gndr,
    year = yrbrn,
    edu_year = eduyrs,
    edu = edulvlb,	
    income = hinctnta,
    region = region
  )
#dput(selected$var_codes)
```

List all variables (old and new names) as well as their range of levels. 

**List of variables for eda**

* `id` (`idno`):
* `cntry` (`cntry`):
* `news_pol` (`nwspol`):
* `inet` (`netusoft`):
* `poli_intr` (`polintr`):
* `have_say` (`psppsgva`):
* `pol_grp` (`actrolga`):
* `have_infl` (`psppipla`):
* `in_eff` 
    * `pol_conf` (`cptppola`):
    * `pol_grp` (`pol_grp`):
* `ex_eff`
    + `trust_parl` (`trstlgl`):
    + `trust_leg` (`trstlgl`)
    + `trust_police` (`trstplc`):
    + `trust_eu` (`trstep`):
    + ~~`trust_un` (`trstun`):~~
* `vote_nat`
    + `vote_at` (`prtvtbat`):
    + `vote_be` (`prtvtcbe`):
    + `vote_ch` (`prtvtfch`):
    + `vote_cz` (`prtvtdcz`):
    + `vote_e1` (`prtvede1`):
    + `vote_fi` (`prtvtdfi`):
    + `vote_fr` (`prtvtcfr`):
    + `vote_gb` (`prtvtbgb`):
    + `vote_ie` (`prtvtbie`):
    + `vote_il` (`prtvtcil`):
    + `vote_is` (`prtvtbis`):
    + `vote_nl` (`prtvtfnl`):
    + `vote_no` (`prtvtbno`):
    + `vote_se` (`prtvtbse`):
    + `vote_si` (`prtvtesi`):
* `party_act` (`wrkprty`):
* `sign_pet` (`sgnptit`):
* `demo` (`pbldmn`):
* `lrscale` (`lrscale`):
* `s_life` (`stflife`):
* `s_econ` (`stfeco`):
* `s_gov` (`stfgov`):
* `s_dem` (`stfdem`):
* `state_edu` (`stfedu`):
* `sate_health` (`stfhlth`):
* `eu_state` (`euftf`):
* `imm`
    + `imm_econ` (`imbgeco`):
    + `imm_same` (`imsmetn`):
    + `imm_diff` (`imdfetn`):
    + `imm_poor` (`impcntr`):
* `rel` (`rlgatnd`):
* `citz` (`ctzcntr`):
* `social_lazy` (`sblazy`):
* `gndr` (`gndr`):
* `year` (`yrbrn`):
* `edu_year` (`eduyrs`):
* `income` (`hinctnta`):
* `region` (`region`):


```{r}
ess_clean <- ess_sub  %>% 
  mutate_if(.predicate = is.double, .funs = as.integer) %>%
  mutate_if(.predicate = is.character, .funs = as.character) %>%
  mutate_if(.predicate = is.factor, .funs = as.factor) %>%
  mutate(eu_member =
           recode_factor(cntry,
                DE = 1958,
                BE = 1958,
                FR = 1958,
                NL = 1958,
                IE = 1973,
                GB = 1973,
                FI = 1995,
                AT = 1995,
                SE = 1995,
                EE = 2004,
                PL = 2004,
                SI = 2004,
                CZ = 2004,
                CH = 0,
                IL = 0,
                IS = 0,
                NO = 0,
                RU = 0
              )
        ) %>%
  mutate(post_com = ifelse(region %in% c("Estonia", "Poland", "Slovenia", "Czech Republic", "Russian Federation"), "Post C", "West"))

glimpse(ess_clean)
length(table(ess_sub$cntry))
```

## Get Info

* better left-right scale more athoritarian
* horse shoe theory
* 


## Data 

* party recoding
* Inspect each variable with a histogram or barplot, check range and validate data. 


### Block 5

* clean Variables
* document range

```{r}
p5 <- ess_clean %>%
  select(gndr, edu, income, rel, year) %>%
  gather("var", "value") %>%
  ggplot(aes(value, fill = var)) +
  geom_bar() +
  facet_wrap(~var, scales = "free") +
  viridis::scale_fill_viridis(discrete = T)
p5
```





## EDA

* Check selective or pairwise compare correlations (stepwiseAIC or multilevel predction tree).
* spatial mapping
* Factor scores (Factor Analysis)
    + Trust (target)
    + Imm (X)

```{r eda1, echo = F, fig.height=6, fig.width=10, results="show"}
# citizens trust in european institutions 
d <- ess_sub

# graph settings
set_theme(#geom.outline.color = "antiquewhite4", 
          #geom.outline.size = 1, 
          #geom.label.size = 2,
          #geom.label.color = "grey50",
          title.color = "black", 
          title.size = 1.5, 
          axis.angle.x = 45, 
          axis.textcolor = "black", 
          base = theme_538())

p1 <- sjp.frq(d$trust_parl, 
        type = "hist", 
        show.mean = TRUE, 
        show.values = FALSE, 
        geom.colors = "#336699", prnt.plot = F)

p2 <- sjp.frq(d$trust_leg, 
        type = "hist", 
        show.mean = TRUE, 
        show.values = FALSE, 
        geom.colors = "#336699", prnt.plot = F)

p3 <- sjp.frq(d$trust_police, 
        type = "hist", 
        show.mean = TRUE, 
        show.values = FALSE, 
        geom.colors = "#336699", prnt.plot = F)

p4 <- sjp.frq(d$trust_parl, 
        type = "hist", 
        show.mean = TRUE, 
        show.values = FALSE, 
        geom.colors = "#336699", prnt.plot = F)

p5 <- sjp.frq(d$trust_party, 
        type = "hist", 
        show.mean = TRUE, 
        show.values = FALSE, 
        geom.colors = "#336699", prnt.plot = F)

p6 <- sjp.frq(d$trust_pol, 
        type = "hist", 
        show.mean = TRUE, 
        show.values = FALSE, 
        geom.colors = "#336699", prnt.plot = F)

library(gridExtra)
grid.arrange(p1$plot, p2$plot, p3$plot, p4$plot, p5$plot, p6$plot, nrow = 3, top = grid::textGrob("Trust in Political Institutions", vjust = 0.5, gp = grid::gpar(fontface = "bold", cex = 2)))
```




## Statistical Modeling

### Multilevel 

* Y = voting right wing/populist
* X_1 = ex_eff (trust factor scores) + in_eff
* X_2 = immigration(factor scores)
* X_3 = s_gov + s_dem + interaction + demo + sign_pet
* X_4 = poli_intr + lrscale + party_act + news_pol
* X_5 = gndr + edu + income + rel + year
* X_6 = social_lazy (social benefit attitudes)
* What about Z predictors?


### Prection Power

* multilevel tree
* `brms`
* viele random slopes
